---
title: "Multi_omics_Analysis"
author: "Deborah,Yann,Brignt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(mixOmics)
library(igraph)
library(org.Hs.eg.db)
library(clusterProfiler)
library(gprofiler2)
library(BiocParallel)
library(netOmics)
library(dplyr)
library(DT)
```
## =============================================================================
## Data Processing
## =============================================================================

```{r}
data <- readRDS("data_hela.Rds")

names(data)
info <- data$sample_info
rna     <- data$mRNA
protein <- data$protein
trans   <- data$trans

cat("Number of samples and genes in the mRNA data ", dim(rna), "\\n")
cat("Number of samples and transcripts ", dim(trans), "\\n")
cat("Number of samples and proteins ", dim(protein), "\\n")
```

```{r}
head(rownames(rna))
summary(as.factor(info$Y))
```

```{r}
boxplot(rna[,1:100],     col = as.factor(info$sample), main = "Boxplot mRNA (first 100 genes)")
boxplot(protein[,1:100], col = as.factor(info$sample), main = "Boxplot protein (first 100 proteins)")
boxplot(trans[,1:100],   col = as.factor(info$sample), main = "Boxplot trans (first 100 transcripts)")
```

```{r}
boxplot(t(rna),     col = as.factor(info$sample), main = "Boxplot mRNA by sample")
boxplot(t(protein), col = as.factor(info$sample), main = "Boxplot protein by sample")
boxplot(t(trans),   col = as.factor(info$sample), main = "Boxplot trans by sample")
```

```{r}
boxplot(t(rna),     col = as.factor(info$Y), main = "Boxplot mRNA by class")
boxplot(t(protein), col = as.factor(info$Y), main = "Boxplot protein by class")
boxplot(t(trans),   col = as.factor(info$Y), main = "Boxplot trans by class")
```

```{r}
coef.var <- function(x) {
  m <- mean(x, na.rm = TRUE)
  s <- sd(x, na.rm = TRUE)
  if (is.na(m) || m == 0) return(NA_real_)
  s / m
}

# Filter mRNA
rna <- rna[, colSums(rna) > 10]
coef.mRNA <- sapply(rna, coef.var)
coef.mRNA <- coef.mRNA[!is.na(coef.mRNA)]
hist(coef.mRNA, main = "Coefficient of variation (mRNA)")

rna.filtered <- rna[, abs(coef.mRNA) > 0.02]
rna.names    <- names(coef.mRNA)[abs(coef.mRNA) > 0.02]
top10_rna    <- names(sort(abs(coef.mRNA), decreasing = TRUE))[1:min(10, length(coef.mRNA))]
```

```{r}
# Filter protein
protein <- protein[, colSums(protein) > 10]
coef.protein <- sapply(protein, coef.var)
coef.protein <- coef.protein[!is.na(coef.protein)]
hist(coef.protein, main = "Coefficient of variation (protein)")

protein.filtered <- protein[, abs(coef.protein) > 0.02]
protein.names    <- names(coef.protein)[abs(coef.protein) > 0.02]
top10_protein    <- names(sort(abs(coef.protein), decreasing = TRUE))[1:min(10, length(coef.protein))]
```

```{r}
# Filter transcripts
trans <- trans[, colSums(trans) > 10]
coef.trans <- sapply(trans, coef.var)
coef.trans <- coef.trans[!is.na(coef.trans)]
hist(coef.trans, main = "Coefficient of variation (trans)")

trans.filtered <- trans[, abs(coef.trans) > 0.02]
trans.names    <- names(coef.trans)[abs(coef.trans) > 0.02]
top10_trans    <- names(sort(abs(coef.trans), decreasing = TRUE))[1:min(10, length(coef.trans))]
```

```{r}
# Scaling
rna.scaled <- scale(rna.filtered, center = TRUE, scale = TRUE)
hist(rna.scaled, main = "Histogram after scaling (mRNA)")

meds <- apply(protein.filtered, 2, median, na.rm = TRUE)
protein.scaled <- sweep(protein.filtered, 2, meds, "-")
protein.scaled <- as.matrix(protein.scaled)
hist(protein.scaled, main = "Histogram after scaling (protein)")

trans.scaled <- scale(trans.filtered, center = TRUE, scale = TRUE)
hist(trans.scaled, main = "Histogram after scaling (trans)")
```

```{r}
boxplot(t(rna.scaled),     col = as.factor(info$sample), main = "Scaled mRNA by sample")
boxplot(t(protein.scaled), col = as.factor(info$sample), main = "Scaled protein by sample")
boxplot(t(trans.scaled),   col = as.factor(info$sample), main = "Scaled trans by sample")

cat("RNA features kept:",     length(rna.names), "\\n")
cat("Protein features kept:", length(protein.names), "\\n")
cat("Trans features kept:",   length(trans.names), "\\n")
top10_rna
top10_protein
top10_trans
```

## =============================================================================
## PCA
## =============================================================================

```{r}
sample <- as.factor(info$sample)  # grouping factor for plots

# PCA for each omics block
pca.rna     <- pca(X = rna.scaled,     ncomp = 5)
pca.trans   <- pca(X = trans.scaled,   ncomp = 5)
pca.protein <- pca(X = protein.scaled, ncomp = 5)

pca.rna
pca.trans
pca.protein

# Sample plots
plotIndiv(pca.rna,     group = sample, legend = TRUE, main = "mRNA PCA")
plotIndiv(pca.trans,   group = sample, legend = TRUE, main = "Transcript PCA")
plotIndiv(pca.protein, group = sample, legend = TRUE, main = "Protein PCA")

# Scree plots
plot(pca.rna)
plot(pca.trans)
plot(pca.protein)
```

```{r}
# Correlation circle plots (variables)
par(mfrow = c(1, 3))
plotVar(pca.rna,     var.names = FALSE, title = "RNA")
plotVar(pca.trans,   var.names = FALSE, title = "Trans")
plotVar(pca.protein, var.names = FALSE, title = "Protein")
par(mfrow = c(1, 1))
```

```{r}
# Top contributing transcript variables to PC1
load_trans <- pca.trans$loadings$X[, 1]
ord        <- order(abs(load_trans), decreasing = TRUE)
top_trans  <- names(load_trans)[ord][1:10]

# g:Profiler enrichment for top PCA transcripts
res.trans.pca <- gost(
  query             = top_trans,
  organism          = "hsapiens",
  correction_method = "fdr",
  evcodes           = TRUE
)

gostplot(res.trans.pca)

res_tab <- res.trans.pca$result
res_tab$p_adj <- p.adjust(res_tab$p_value, method = "fdr")

res_trans_pca_simple <- res_tab[, c(
  "source",
  "term_name",
  "p_value",
  "p_adj",
  "term_size",
  "query_size",
  "intersection_size",
  "intersection"
)]

head(res_trans_pca_simple)
```

```{r}
# Example loading plot for mRNA (component 2)
plotLoadings(
  pca.rna,
  comp     = 2,
  block    = "Y",
  ndisplay = 10,
  title    = "Top mRNA loadings, component 2"
)
```

## =============================================================================
## sPCA
## =============================================================================

```{r}
# sPCA on mRNA
pca.rna.sp <- spca(rna.scaled, ncomp = 2, keepX = c(10, 5))
plotVar(pca.rna.sp, var.names = TRUE, title = "sPCA mRNA variables")

plotLoadings(
  pca.rna.sp,
  comp     = 2,
  block    = "Y",
  ndisplay = 10,
  title    = "Top mRNA loadings (sPCA), component 2"
)
```

```{r}
# sPCA on transcripts
pca.trans.sp <- spca(trans.scaled, ncomp = 2, keepX = c(10, 5))
plotVar(pca.trans.sp, var.names = TRUE, title = "sPCA trans variables")
```

```{r}
# Enrichment for sPCA-selected transcript variables (component 1)
trans_genes_c1 <- selectVar(pca.trans.sp, comp = 1)$name

res.trans.c1 <- gost(
  query             = trans_genes_c1,
  organism          = "hsapiens",
  correction_method = "fdr",
  evcodes           = TRUE
)

gostplot(res.trans.c1)

res_tab <- res.trans.c1$result
res_tab$p_adj <- p.adjust(res_tab$p_value, method = "fdr")

res_trans_simple <- res_tab[, c(
  "source",
  "term_name",
  "p_value",
  "p_adj",
  "term_size",
  "query_size",
  "intersection_size",
  "intersection"
)]

head(res_trans_simple)
```

```{r}
# sPCA on protein
pca.protein.sp <- spca(protein.scaled, ncomp = 2, keepX = c(10, 5))
plotVar(pca.protein.sp, var.names = TRUE, title = "sPCA protein variables")
```

## =============================================================================
## SUPERVISED ANALYSIS (PLS-DA)
## =============================================================================

```{r}
# PLS-DA on mRNA
group  <- info$Y

plsda.rna <- plsda(rna.scaled, group, ncomp = 2)
plotIndiv(plsda.rna, group = group, legend = TRUE, title = "PLS-DA mRNA")
```

```{r}
# PLS-DA on transcripts
plsda.trans <- plsda(trans.scaled, sample, ncomp = 2)
plotIndiv(plsda.trans, group = sample, legend = TRUE, title = "PLS-DA trans")
```
```{r}
# PLS-DA on proteins
plsda.protein <- plsda(protein.scaled, sample, ncomp = 2)
plotIndiv(plsda.protein, group = sample, legend = TRUE, title = "Protein PLS-DA")
```

## =============================================================================
## PLS
## =============================================================================

```{r}
# PLS to model associations between mRNA (X) and protein (Y)
pls.res <- pls(X = rna.scaled, Y = protein.scaled, ncomp = 2)

# Cross-validation to evaluate performance
perf.pls <- perf(
  pls.res,
  validation  = "Mfold",
  folds       = 5,
  nrepeat     = 10,
  progressBar = TRUE,
  seed        = 24
)

# Performance metrics
#perf.pls$measure$MSEP
#perf.pls$measure$R2
#perf.pls$measure$Q2.total
```

```{r}
# PLS visualisations
plotIndiv(pls.res, comp = c(1, 2), group = sample,
          legend = TRUE, main = "PLS mRNA–protein (samples)")

plotVar(pls.res, comp = c(1, 2), var.names = FALSE,legend = TRUE)

plotArrow(pls.res, main = "Arrow plot: correlation between X and Y scores")
```

## =============================================================================
## sPLS
## =============================================================================

```{r}
# Sparse PLS for RNA–protein integration with feature selection
spls.res <- spls(
  X      = rna.scaled,
  Y      = protein.scaled,
  ncomp  = 2,
  keepX  = c(10, 5),
  keepY  = c(10, 5)
)

# Selected variables on each component
selectVar(spls.res, comp = 1)$X$name
selectVar(spls.res, comp = 1)$Y$name
selectVar(spls.res, comp = 2)$X$name
selectVar(spls.res, comp = 2)$Y$name
```

```{r}
# Loadings for RNA block (component 1)
plotLoadings(
  spls.res,
  comp     = 1,
  block    = "X",
  ndisplay = 20,
  title    = "Top RNA loadings, sPLS component 1"
)

# Loadings for protein block (component 1)
plotLoadings(
  spls.res,
  comp     = 1,
  block    = "Y",
  ndisplay = 20,
  title    = "Top protein loadings, sPLS component 1"
)
```

```{r}
# Correlation circle plot for sPLS variables
plotVar(
  spls.res,
  comp      = c(1, 2),
  var.names = FALSE,
  title     = "sPLS variable correlation plot"
)
```

```{r}
# Sample plot for sPLS (RNA–protein)
plotIndiv(
  spls.res,
  ind.names = TRUE,
  group     = sample,
  legend    = TRUE,
  main      = "sPLS sample plot (RNA–protein)"
)
```

```{r}
# Enrichment of sPLS component 1 RNA features
genes1    <- selectVar(spls.res, comp = 1)$X$name
proteins1 <- selectVar(spls.res, comp = 1)$Y$name

res.genes1 <- gost(
  query             = genes1,
  organism          = "hsapiens",
  correction_method = "fdr",
  evcodes           = TRUE,
  multi_query       = FALSE
)

gostplot(res.genes1)

res_tab <- res.genes1$result
res_genes_simple <- res_tab[, c(
  "source",
  "term_name",
  "p_value",
  "term_size",
  "query_size",
  "intersection_size",
  "intersection"
)]

head(res_genes_simple)

datatable(
  res_genes_simple,
  options = list(pageLength = 20),
  caption = "Enriched terms for genes1 with genes involved"
)
```

```{r}
# Enrichment of sPLS component 1 protein features
res.protein1 <- gost(
  query             = proteins1,
  organism          = "hsapiens",
  correction_method = "fdr",
  evcodes           = TRUE,
  multi_query       = FALSE
)

gostplot(res.protein1)

res_tab <- res.protein1$result
res_protein_simple <- res_tab[, c(
  "source",
  "term_name",
  "p_value",
  "term_size",
  "query_size",
  "intersection_size",
  "intersection"
)]

head(res_protein_simple)

datatable(
  res_protein_simple,
  options = list(pageLength = 5),
  caption = "Enriched terms for proteins1 with genes involved"
)
```

```{r}
#  Enrichment for component 2 features
genes2    <- selectVar(spls.res, comp = 2)$X$name
proteins2 <- selectVar(spls.res, comp = 2)$Y$name

res.genes2 <- gost(
  query             = genes2,
  organism          = "hsapiens",
  correction_method = "fdr"
)
gostplot(res.genes2)

res.protein2 <- gost(
  query             = proteins2,
  organism          = "hsapiens",
  correction_method = "fdr"
)
gostplot(res.protein2)
```

## =============================================================================
## DIABLO
## =============================================================================

```{r}
# Align samples across omics and metadata
common.samples <- Reduce(intersect, list(
  rownames(rna.scaled),
  rownames(protein.scaled),
  rownames(trans.scaled),
  info$sample
))

rna.scaled     <- rna.scaled[common.samples, , drop = FALSE]
protein.scaled <- protein.scaled[common.samples, , drop = FALSE]
trans.scaled   <- trans.scaled[common.samples, , drop = FALSE]
info           <- info[match(common.samples, info$sample), ]

stopifnot(all(rownames(rna.scaled)     == common.samples))
stopifnot(all(rownames(protein.scaled) == common.samples))
stopifnot(all(rownames(trans.scaled)   == common.samples))
stopifnot(all(info$sample              == common.samples))

Y <- factor(info$Y)

sapply(list(rna = rna.scaled,
            protein = protein.scaled,
            trans = trans.scaled), dim)
length(Y)
```

```{r}
## 2-block DIABLO: rna + protein
X_2 <- list(
  rna     = rna.scaled,
  protein = protein.scaled
)

design_2 <- matrix(0.1, nrow = length(X_2), ncol = length(X_2))
diag(design_2) <- 0
colnames(design_2) <- rownames(design_2) <- names(X_2)

test.keepX_2 <- list(
  rna     = c(5, 10),
  protein = c(5, 10)
)

set.seed(123)
tune_2 <- tune.block.splsda(
  X          = X_2,
  Y          = Y,
  ncomp      = 2,
  test.keepX = test.keepX_2,
  design     = design_2,
  validation = "Mfold",
  folds      = 3,
  nrepeat    =10,
  dist       = "centroids.dist",
  measure    = "BER",
  progressBar = TRUE,
  BPPARAM     = SerialParam()
)

final.keepX_2 <- tune_2$choice.keepX

diablo_2 <- block.splsda(
  X      = X_2,
  Y      = Y,
  ncomp  = 2,
  keepX  = final.keepX_2,
  design = design_2
)

perf_2 <- perf(
  diablo_2,
  validation = "Mfold",
  folds      = 3,
  nrepeat    = 10
)
```


```{r}
## Overall object
perf_2

## 1) Balanced error rate (BER) per component
perf_2$error.rate  # or $error.rate$overall depending on version
perf_2$error.rate.per.class.all

## 2) Plot error vs number of components
plot(perf_2)    # built-in plot: error rate per component

```

```{r}
# 2-block DIABLO plots to PDF
#png("diablo_genes_rna1.png", width = 2400, height = 1800, res = 300)
plotDiablo(diablo_2, ncomp = 1)
circosPlot(diablo_2, comp = 1, cutoff = 0.6, size.variables = 0.5)
#dev.off()

#png("diablo_genes_rna2.png", width = 2400, height = 1800, res = 300)

plotDiablo(diablo_2, ncomp = 2)
circosPlot(diablo_2, comp = 2, cutoff = 0.6, size.variables = 0.5)

#dev.off()

comp_list1 <- list(rna = 1, protein = 1)
comp_list2 <- list(rna = 2, protein = 2)

#png("Network_genes_rna1.png", width = 2400, height = 1800, res = 300)
#par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))
#network(diablo_2, comp = comp_list1, cutoff = 0.6,
       # graph.scale = 0.4, cex.node.name = 0.5)
#dev.off()

#png("Network_genes_rna2.png", width = 2400, height = 1800, res = 300)
#par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))
#network(diablo_2, comp = comp_list2, cutoff = 0.6,
        #graph.scale = 0.4, cex.node.name = 0.5)
#dev.off()

plotVar(diablo_2, block = "rna",     comp = c(1, 2), var.names = FALSE)
plotVar(diablo_2, block = "protein", comp = c(1, 2), var.names = FALSE)
```

```{r}
## 3-block DIABLO: rna + protein + trans
X_3 <- list(
  rna     = rna.scaled,
  protein = protein.scaled,
  trans   = trans.scaled
)

design_3 <- matrix(0.1, nrow = length(X_3), ncol = length(X_3))
diag(design_3) <- 0
colnames(design_3) <- rownames(design_3) <- names(X_3)

test.keepX_3 <- list(
  rna     = c(5, 10),
  protein = c(5, 10),
  trans   = c(5, 10)
)

set.seed(123)
tune_3 <- tune.block.splsda(
  X          = X_3,
  Y          = Y,
  ncomp      = 3,
  test.keepX = test.keepX_3,
  design     = design_3,
  validation = "Mfold",
  folds      = 3,
  nrepeat    = 10,
  dist       = "centroids.dist",
  measure    = "BER",
  progressBar = TRUE,
  BPPARAM     = BiocParallel::SerialParam()
)

final.keepX_3 <- tune_3$choice.keepX

diablo_3 <- block.splsda(
  X      = X_3,
  Y      = Y,
  ncomp  = 3,
  keepX  = final.keepX_3,
  design = design_3
)

names(diablo_3$X)

#png("net_comp1_proteins_genes_trans.png", width = 2400, height = 1800, res = 300)
plotDiablo(diablo_3, ncomp = 1)
circosPlot(diablo_3, comp = 1, cutoff = 0.6, size.variables = 0.5)
#dev.off()

comp_list1 <- list(rna = 1, protein = 1)
#png("Network_genes_rna_trans1.png", width = 2400, height = 1800, res = 300)
#par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))
#network(diablo_3, comp = comp_list1, cutoff = 0.6,
        #graph.scale = 0.4, cex.node.name = 0.5)
#dev.off()


#png("net_comp2_proteins_genes_trans.png", width = 2400, height = 1800, res = 300)
plotDiablo(diablo_3, ncomp = 2)
circosPlot(diablo_3, comp = 2, cutoff = 0.6, size.variables = 0.5)
#dev.off()
```

```{r}
## Performance of 3-block DIABLO
perf_3 <- perf(
  diablo_3,
  validation = "Mfold",
  folds      = 3,
  nrepeat    = 10
)

perf_3          

## 2) Plot error vs number of components
plot(perf_3)

```

##==============================================================================
## NETWORK ANALYSIS
##==============================================================================


```{r}
################################################################################
# STEP 7 - NETWORK ANALYSIS
# Part 1: Gene Regulatory Network (GRN)
################################################################################

genes.use <- unique(c(
  selectVar(diablo_2, block = "rna", comp = 1)$rna$name,
  selectVar(diablo_2, block = "rna", comp = 2)$rna$name
))

proteins.use <- unique(c(
  selectVar(diablo_2, block = "protein", comp = 1)$protein$name,
  selectVar(diablo_2, block = "protein", comp = 2)$protein$name
))

genes.use
proteins.use
length(genes.use)
length(proteins.use)
```

```{r}

library(igraph)

# Subset RNA matrix on selected DIABLO genes
rna.net <- rna.scaled[, colnames(rna.scaled) %in% genes.use, drop = FALSE]

dim(rna.net)
colnames(rna.net)

# Correlation matrix between selected genes
cor.grn <- cor(rna.net, use = "pairwise.complete.obs", method = "pearson")

# Threshold for edge creation
thr.grn <- 0.7

adj.grn <- abs(cor.grn) >= thr.grn
diag(adj.grn) <- 0

# Build graph
g.grn <- graph_from_adjacency_matrix(adj.grn, mode = "undirected", diag = FALSE)
g.grn <- igraph::simplify(g.grn, remove.multiple = TRUE, remove.loops = TRUE)

# Basic statistics
grn_nodes <- vcount(g.grn)
grn_edges <- ecount(g.grn)
grn_degree <- degree(g.grn)
grn_isolated <- sum(grn_degree == 0)
grn_top_gene <- names(which.max(grn_degree))

grn_nodes
grn_edges
grn_isolated
grn_top_gene

sort(grn_degree, decreasing = TRUE)
```


```{r}
hist(grn_degree,
     main = "GRN degree distribution",
     xlab = "Degree",
     col = "lightblue",
     border = "black")


set.seed(1)
lay <- layout_with_kk(g.grn)   # Fruchterman–Reingold (more spread)
# alternatives: layout_with_kk, layout_with_graphopt, layout_with_drl

png("GRN_big_nodes_spread.png", width = 1600, height = 1200, res = 200)

plot(
  g.grn,
  layout             = lay,
  main               = "Gene Regulatory Network",
  vertex.size        = 45,
  vertex.label.cex   = 0.45,
  vertex.label.color = "black"
)

dev.off()
```

```{r}
grn_density <- igraph::edge_density(g.grn)
grn_components <- igraph::components(g.grn)

grn_density
grn_components$no
grn_components$csize
```
```{r}
grn_summary <- data.frame(
  nodes = grn_nodes,
  edges = grn_edges,
  isolated_nodes = grn_isolated,
  density = grn_density,
  connected_components = grn_components$no,
  top_gene = grn_top_gene
)

sort(grn_degree, decreasing = TRUE)
```

```{r}
human_gene_protein <- readRDS("human_gene_to_coding_protein.Rds")
human_ppi <- readRDS("human_ppi.Rds")
human_tf_gene <- readRDS("human_TF_to_targeted_gene.Rds")
```


```{r}
class(human_gene_protein)
class(human_ppi)
class(human_tf_gene)

str(human_gene_protein)
str(human_ppi)
str(human_tf_gene)

head(human_gene_protein)
head(human_ppi)
head(human_tf_gene)

colnames(human_gene_protein)
colnames(human_ppi)
colnames(human_tf_gene)
```
```{r}

gp <- as.data.frame(human_gene_protein)
ppi <- as.data.frame(human_ppi)
tfg <- as.data.frame(human_tf_gene)

gp <- gp[, 1:2, drop = FALSE]
ppi <- ppi[, 1:2, drop = FALSE]
tfg <- tfg[, 1:2, drop = FALSE]

colnames(gp) <- c("from", "to")
colnames(ppi) <- c("from", "to")
colnames(tfg) <- c("from", "to")

head(gp)
head(ppi)
head(tfg)
```


```{r}
#PPI
ppi.sub <- ppi %>%
  filter(from %in% proteins.use, to %in% proteins.use)

g.ppi <- graph_from_data_frame(ppi.sub, directed = FALSE)
g.ppi <- igraph::simplify(g.ppi, remove.multiple = TRUE, remove.loops = TRUE)

ppi_nodes <- igraph::vcount(g.ppi)
ppi_edges <- igraph::ecount(g.ppi)
ppi_degree <- igraph::degree(g.ppi)
ppi_isolated <- sum(ppi_degree == 0)
ppi_top_protein <- names(which.max(ppi_degree))

ppi_nodes
ppi_edges
ppi_isolated
ppi_top_protein
sort(ppi_degree, decreasing = TRUE)
```

```{r}
#PROTEIN CODINNG LINKS
gp.sub <- gp %>%
  filter(from %in% genes.use, to %in% proteins.use)

gp.sub
nrow(gp.sub)
```

```{r}
#TF-GENE LINKS
tfg.sub <- tfg %>%
  filter(to %in% genes.use)
nrow(tfg.sub)
head(tfg.sub)
```

```{r}
#TF-GENE RESTREINT TO SELECTED PROTEINS
tfg.sub2 <- tfg %>%
  filter(from %in% proteins.use, to %in% genes.use)

nrow(tfg.sub2)
head(tfg.sub2)
```
```{r}
# EXTENDED PPI
ppi.sub.ext <- ppi %>%
  filter(from %in% proteins.use | to %in% proteins.use)

head(ppi.sub.ext)
nrow(ppi.sub.ext)

g.ppi.ext <- graph_from_data_frame(ppi.sub.ext, directed = FALSE)
g.ppi.ext <- igraph::simplify(g.ppi.ext, remove.multiple = TRUE, remove.loops = TRUE)

ppi_ext_nodes <- igraph::vcount(g.ppi.ext)
ppi_ext_edges <- igraph::ecount(g.ppi.ext)
ppi_ext_degree <- igraph::degree(g.ppi.ext)
ppi_ext_top <- names(which.max(ppi_ext_degree))

ppi_ext_nodes
ppi_ext_edges
ppi_ext_top
sort(ppi_ext_degree, decreasing = TRUE)[1:10]
```
The combined network is dominated by protein hubs, with P46783 emerging as the
most central node, indicating a potential key regulatory or structural role in 
the integrated system.
## =============================================================================
## MULTI OMICS NETWORK
## =============================================================================
```{r}
# COMBINED MULTI-OMICS NETWORK

# GRN edges
edges.grn <- as_data_frame(g.grn, what = "edges")
edges.grn$type <- "GRN"

# Extended PPI edges
edges.ppi <- as_data_frame(g.ppi.ext, what = "edges")
edges.ppi$type <- "PPI"

# TF -> gene edges
edges.tfg <- tfg.sub
edges.tfg$type <- "TF_gene"

# Combine all edges
all.edges <- bind_rows(edges.grn, edges.ppi, edges.tfg) %>%
  distinct()

head(all.edges)
nrow(all.edges)

# Build combined network
g.multi <- graph_from_data_frame(all.edges, directed = FALSE)
g.multi <- igraph::simplify(g.multi, remove.multiple = TRUE, remove.loops = TRUE)

# Basic stats
multi_nodes <- igraph::vcount(g.multi)
multi_edges <- igraph::ecount(g.multi)
multi_degree <- igraph::degree(g.multi)
multi_density <- igraph::edge_density(g.multi)
multi_components <- igraph::components(g.multi)

multi_nodes
multi_edges
multi_density
multi_components$no
multi_components$csize
sort(multi_degree, decreasing = TRUE)[1:10]
```

## ============================================================================
## MODULARITY ANALYSIS
## ============================================================================
```{r}
clust <- cluster_louvain(g.multi)

mod_score <- modularity(clust)
module_membership <- membership(clust)

mod_score
table(module_membership)
```


```{r}

# 1) Define which nodes to label (example: top 20 by degree)
deg <- degree(g.multi)
top_ids <- order(deg, decreasing = TRUE)[1:10]

labels <- rep(NA, vcount(g.multi))
labels[top_ids] <- V(g.multi)$name[top_ids]


set.seed(1)
lay <- layout_with_fr(g.multi)   

# 2) PNG device
pdf("multiomics_network_big.pdf", width = 12, height = 10)
plot(
  clust, g.multi,
  layout           = lay,
  vertex.size      = 12,
  vertex.label     = labels,
  vertex.label.cex = 0.5,
  main             = "Combined multi-omics network with Louvain modules"
)
dev.off()
```
The modularity score (~0.62) indicates a strong community structure, suggesting 
that the network is organized into well-defined functional modules.
The network is structured into six modules of varying sizes, with one dominant 
module and several smaller specialized communities

## ============================================================================
## RANDOM WALK
## ============================================================================
```{r}
seed.node <- names(which.max(multi_degree))
seed.node
set.seed(123)
rw <- random_walk(g.multi, start = seed.node, steps = 50)
rw_freq <- sort(table(rw), decreasing = TRUE)
rw_freq
head(rw_freq, 10)
```

```{r}
module_sizes <- sort(table(module_membership), decreasing = TRUE)
module_sizes

largest_module <- as.numeric(names(module_sizes)[1])
largest_nodes <- names(module_membership[module_membership == largest_module])


length(largest_nodes)

top_module_nodes <- names(sort(multi_degree[largest_nodes], decreasing = TRUE))[1:50]

res_module <- gost(query = top_module_nodes, organism = "hsapiens",correction_method = "fdr",
  evcodes           = TRUE,
  multi_query       = FALSE)

res.tab <- res_module$result
gostplot(res_module)
res.table  <- res.tab[, c(
  "source",
  "term_name",
  "p_value",
  "term_size",
  "query_size",
  "intersection_size",
  "intersection"
)]

datatable(res.table
          ,options = list(pageLength = 20),
  caption = "Enriched terms for involved")

```
The largest module of the integrated network showed strong enrichment for CORUM 
complexes, suggesting that the nodes grouped by the modularity analysis 
participate in shared molecular assemblies. This reinforces the biological 
validity of the network structure and indicates that the integrated graph 
captures coordinated functional units.

## CONCLUSION

A multi-mics network was constructed by integrating the gene regulatory network, 
the extended protein-protein interaction network, and TF-gene relationships. 
The final graph contained 1546 nodes and 783 edges and formed a single connected 
component, showing that the different omics layers were successfully integrated 
into a coherent network. Topological analysis revealed a sparse but highly 
structured biological network, dominated by central hub proteins such as P46783. 
Louvain clustering identified six modules with a high modularity score (0.6169), 
indicating strong community organization. A random walk initiated from the main 
hub remained centered around this node, confirming its key topological role. 
Finally, enrichment analysis of the largest module identified several significantly 
enriched CORUM complexes, supporting the biological relevance of the detected 
communities.


